### 2021-01-28 letter.tex
:memo: Initialize response letter

\documentclass[a4paper, 10pt]{letter}

\name{Yang Zhao, Bruno Clerckx and Zhenyuan Feng}
\date{}

\begin{document}
	\begin{letter}{Prof. Liang Xiao \\ Editor \\ IEEE Transactions on Communications \\ ~ \\ Re: Decision on TCOM-TPS-20-1501 (Intelligent Reflecting Surface-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design)}
		\opening{Dear Prof. Liang Xiao and Reviewers,}

		We would like to express our appreciation for the time and effort dedicated to the reviewing of our paper. The manuscript has been revised carefully based on your valuable comments and suggestions. In particular, we attempted to clarify the motivations and contributions of this work. To answer the questions and concerns, we include a point-to-point response below, which also describes all modifications made to the manuscript. We hope that the revisions are to the satisfaction of the editor and reviewers.

		\closing{Best Regards,}
	\end{letter}
\end{document}