### 2021-04-22 letter.tex

\documentclass[a4paper, 10pt]{letter}

\name{Yang Zhao, Bruno Clerckx and Zhenyuan Feng}
\date{}

\begin{document}
	\begin{letter}{Prof. Liang Xiao \\ Editor \\ IEEE Transactions on Communications \\ ~ \\ Re: Decision on TCOM-TPS-20-1501 (IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design Under Nonlinear Harvester Model)}
		\opening{Dear Prof. Liang Xiao and Reviewers,}

		We would like to express our appreciation for the time and effort dedicated to the reviewing of our paper. The manuscript has been revised carefully based on your valuable comments and suggestions. In particular, we clarified the motivations and contributions of this work, proved the unit-rank property of the obtained IRS matrix, added two low-complexity designs with near-optimal rate-energy performance, and considered the cases of imperfect cascaded CSIT and finite IRS reflection states. To answer the questions and concerns, we include a point-to-point response below, which also describes all modifications made to the manuscript. We hope that the revisions are to the satisfaction of the editor and reviewers.

		\closing{Best Regards,}
	\end{letter}
\end{document}