### 2021-08-12 review/letter_2.tex
:package: Add review-related files

\documentclass[a4paper, 10pt]{letter}

\name{Yang Zhao, Bruno Clerckx and Zhenyuan Feng}
\date{}

\begin{document}
	\begin{letter}{Prof. Liang Xiao \\ Editor \\ IEEE Transactions on Communications \\ ~ \\ Re: Decision on TCOM-TPS-21-0553 (IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design Under Nonlinear Harvester Model)}
		\opening{Dear Prof. Liang Xiao and Reviewers,}

		We would like to express our appreciation for the time and effort dedicated to the reviewing of our paper. The manuscript has been revised carefully based on your valuable comments and suggestions. In particular, we added computational complexity analysis for all proposed algorithms, addressed the optimality issue regarding Algorithm 4, discussed the convergence conditions in Algorithm 5, emphasized the design challenge introduced by passive beamforming and nonlinear EH model, and updated the simulation parameters and reference list. To answer the questions and concerns, we also include a point-to-point response below to discuss all modifications made to the manuscript. We hope that the revisions are to the satisfaction of the editor and reviewers.

		\closing{Best Regards,}
	\end{letter}
\end{document}