### 2021-09-16 review/letter_3.tex

\documentclass[a4paper, 10pt]{letter}

\name{Yang Zhao, Bruno Clerckx and Zhenyuan Feng}
\date{}

\begin{document}
	\begin{letter}{Prof. Liang Xiao \\ Editor \\ IEEE Transactions on Communications \\ ~ \\ Re: Decision on TCOM-TPS-21-0553.R1 (IRS-Aided SWIPT: Joint Waveform, Active and Passive Beamforming Design Under Nonlinear Harvester Model)}
		\opening{Dear Prof. Liang Xiao and Reviewers,}

		We would like to express our appreciation for the time and effort dedicated to the reviewing of our paper. The manuscript has been revised carefully based on your valuable comments and suggestions. We hope that the revisions are to the satisfaction of the editor and reviewers.

		\closing{Best Regards,}
	\end{letter}
\end{document}