### 2021-11-16 assets/ps_receiver.tex
:bento: Update assets

\begin{circuitikz}[transform shape]
	\draw (0,0) node[bareRXantenna](r){Rx}(r.center)
		to[short,-o] ++(0,-1)
		to node[coupler,anchor=left up](c){Power Splitter} ++(1,0);
	\draw (c.right up)
		to[short,-o] ++(0.5,0) node[above right]{$1-\rho$}
		to ++(0,0.5)
		to ++(1,0)
		to node[draw,minimum height=2em,minimum width=10em,anchor=west]{Information Decoder} ++(0,0);
	\draw (c.right down)
		to[short,-o] ++(0.5,0) node[below right]{$\rho$}
		to ++(0,-0.5)
		to ++(1,0)
		to node[draw,minimum height=2em,minimum width=10em,anchor=west]{Energy Harvester} ++(0,0);
\end{circuitikz}